Dynamic Load Balancing for\\Heterogeneous Systems

### 2011-12-05 paper/paper.tex
Reworked Load Balancing Benchmarks.

\section{Load Balancing Benchmarks}
For this paper, three different benchmarks were used to assess the
effectiveness of the load balancer.  The first is a reduction benchmark,
in which the sum of variably-sized array is computed.  The reduction benchmark
is unique in that different code is being run on the CPU and the GPU.  The
reason for this is that the GPU cannot efficiently compute a reduction the way
a CPU typically would -- due to the relative inefficiency of backward branches
on the GPU.  A reduction on the GPU is a multi-stage process, in which the
output of one step is used as the input of the next.  Each steam processor
computes the reduction of two elements during each iteration, reducing the amount
of work during each step by a factor of two.  As a result, each step utilizes half
the stream processors of the previous step.  This continues until the reduction
operation is only done by a single stream processor.  A natural consequence of
this algorithm is that not all of the stream processors will be utilized during
the computation.

The next benchmark is a simple vector addition operation that computes the sum
of two input vectors, and stores the result into an output vector.  The third
benchmark, referred henceforth as the ``Vector Add+'' benchmark, is similar to
the Vector Add benchmark, but it computes 1000 additions before storing the result
in the output vector.  The motivation behind this benchmark is to reduce memory
access overhead with more computation.  All numbers reported in the results section
are given as a mean of ten runs for that particular benchmark, load-balancing scheme,
and data size.

>>>>

\section{Load Balancing Benchmarks}
For this paper, three different benchmarks were used to assess the
effectiveness of the load balancer.  These benchmarks were chosen to represent
tasks that would actually be performed on a GPU, and have seen use in already
existing benchmark suites~\cite{Danalis2010}.  The applications that we have chosen
for this project are VectorAdd, VectorAdd+, and Reduce.

The VectorAdd application is a simple vector addition operation that computes the sum
of two input vectors, and stores the result into an output vector.  This represents the
vector operation $\vec{C} = \vec{A} + \vec{B}$.  This kernel is naive to parallelize, as
it exhibits embarrassingly parallel structure.  Each thread being run maps to a single 
index in the array and computes the output value for that index.  This algorithm is also
nicely designed for load-balancing because of the lack of data dependencies, reducing 
the amount of communication to the GPU.

The next benchmark, referred henceforth as the ``Vector Add+'' benchmark, is similar to
the VectorAdd benchmark, but it computes 1000 additions before storing the result
in the output vector.  The motivation behind this benchmark is to reduce memory
access overhead with more computation.  This benchmark was compiled without optimizations
so that the additions would not be optmized into a single multiplication.  The data
dependencies are exactly the same as for the VectorAdd application, so the results for
data transfer speeds should be equal for both applications, and the difference between
the two should only be the kernel execution times.

The Reduce application is an application that performs a reduction operation on a
large array of data.  In this case, the sum of a variably-sized array is computed.
The Reduce benchmark is unique in that different code is being run on the CPU and 
the GPU.  The reason for this is that the GPU cannot efficiently compute a reduction
the way a CPU typically would -- due to the relatively slower cores and much greater
memory latency.  A optimized reduction on the GPU is a multi-stage process, in which the
output of one step is used as the input of the next.  Each SM computes the reduction
of a smaller set of elements from the larger array during each iteration.  This
requires the use of numerous local atomic operations, as well as local memory to 
facilitate optimized memory access between threads on an SM.
This process continues until the reduction operation is only done by a single SM.
A natural consequence of this algorithm is that not all of the SMs will be utilized during
the computation.  In addition, inside an iteration all of the streaming processors
will only be used for the first stage of reduction, greatly reducing utilization on the
device.

All numbers reported in the results section
are given as a mean of ten runs for that particular benchmark, load-balancing scheme,
and data size.
